---
title: "report14"
author: "Haoliang Zheng"
date: "2/19/2022"
output:
  pdf_document:
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "/Users/zhl/Desktop/drdisc")
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.align = 'center', 
                      fig.width=6, fig.height=7)

library(abind)
library(knitr)
library(latex2exp)
require(magrittr)
require(plyr)
library(tidyverse)
library(coda)
library(doParallel)
library(doMC) 
library(foreach)
library(doRNG)

adapt <- FALSE
bunch <- FALSE
N_para <- 6

registerDoMC(6)

par(mar = c(2,2,2,2))
```


```{r source, include=FALSE}
source("./codes/new-codes.R")
source("./codes/test-codes.R")
source("./codes/other-codes.R")

multi_chain_coda_noa <- function(result_multi, burn = 1000, N_para, p, order){
  coda_result <- list()
  matrix_result <- NULL
  for(i in 1:N_para){
    matrix_result <- NULL
    for(j in 1:order){
      for(k in 1:p){
        bijk <- matrix(result_multi[[i]]$b[j,k,], ncol = 1)
        colnames(bijk) <- paste0("b",j,k)
        matrix_result <- cbind(matrix_result, bijk)
      }
    }
    coda_result[[i]] <- mcmc(matrix_result, start = burn+1)
  }
  coda_result_list <- mcmc.list(coda_result)
  
  return(coda_result_list)
}

para_temp_plot <- function(adapt_PT_new, L, N_para = 6){
par(mar = c(2,2,2,2))
par(mfrow = c(N_para, 1))
for(i in 1:N_para){
  plot(adapt_PT_new[[i]]$temp[1,], type = "l", ylim=range(adapt_PT_new[[i]]$temp),
       main = paste0("starting point ", i))
  for(j in 2:L){
    lines(adapt_PT_new[[i]]$temp[j,], col=tcol(j,.5), lwd=2)
    }
  }
}
```

# math

Denote $T_i$ as inverse temperature i, then the tempered version model is

$$
f_i(y|x) = \frac{\Phi \big\{T_i \big(\sum_{k=1}^K (x^T\beta_k) P_k(y) - (x^T \alpha)_{+} Q(y; \lambda) I(y > 0)\big)\big\}}{\int_{-1}^{1} \Phi \big\{T_i \big(\sum_{k=1}^K (x^T\beta_k) P_k(y) - (x^T \alpha)_{+} Q(y; \lambda) I(y > 0)\big)\big\} \mathrm{d}t}, 
\quad x \in \mathbb{R}^p, y \in (-1, 1),
$$

(
I also tried:
$$
f_i(y|x) = \frac{\Phi \big(\sum_{k=1}^K (x^T\beta_k) P_k(y) - T_i * (x^T \alpha)_{+} Q(y; \lambda) I(y > 0)\big)\big\}}{\int_{-1}^{1} \Phi \big(\sum_{k=1}^K (x^T\beta_k) P_k(y) - T_i * (x^T \alpha)_{+} Q(y; \lambda) I(y > 0)\big) \mathrm{d}t}, 
\quad x \in \mathbb{R}^p, y \in (-1, 1),
$$
)

To simplify the expression, denote the posterior distribution under $T_i$ as:
$$
\pi_i(a) = p_i(x | a) p(a) / c_i
$$

The acceptance probability for inverse temperature moving from 1 to 2 is:
$$
\alpha(1, 2) = 1 \wedge \frac{p_2(x | a) / c_2}{p_1(x | a) / c_1}
$$

Then swap probability is:
$$
\begin{aligned}
\alpha(1, 2) &= 1 \wedge \frac{p_2(x | a_1) / c_2}{p_1(x | a_1) / c_1}
\times \frac{p_1(x | a_2) / c_1}{p_2(x | a_2) / c_2} \\
&= 1 \wedge \frac{p_1(x | a_2) * p_2(x | a_1)}{p_1(x | a_1) * p_2(x | a_2)}
\end{aligned}
$$

# code

## temp

```{r eval=FALSE}
w <- temp * (rowSums(x.remain * (pm %*% b)) + xa.remain*hk)
u <- (runif(n.remain) < Phi(w))
w.miss <- c(w.miss, w[!u])

w.obs <- temp * (rowSums(x * (Poly.obs %*% b)) + xa.obs * Half.obs)

## update b
pg.residual <- pg.residual + temp * c(xPoly.j %*% b[,j])
pg.residual <- pg.residual - temp * c(xPoly.j %*% b[,j])

## update a
pg.residual <- pg.residual + temp * xa * Half.comb
res.new <- pg.residual - temp * xa.new * Half.comb
res <- pg.residual - temp * xa * Half.comb
log.hastings.ratio <- ( sum(dnorm(res.new,0,sqrt(1/pg.draws),log=TRUE)
                            - dnorm(res,0,sqrt(1/pg.draws),log=TRUE))
                        + sum(dnorm(a.new,0,sqrt(1/hprec),log=TRUE)
                              - dnorm(a,0,sqrt(1/hprec),log=TRUE)) )
```

## likelihood ratio

```{r eval=FALSE}
get.swap.likelihood <-  function(temp, 
                                 pg.residual_old, x.comb_old, Half.comb_old,
                                 pg.draws_old, Poly.comb_old,
                                 b_old, b_new,
                                 a_old, a_new, positive=TRUE){
  ## update b
  pg.residual_old <- pg.residual_old + temp * rowSums(x.comb_old * (Poly.comb_old %*% b_old))
  pg.residual_old <- pg.residual_old - temp * rowSums(x.comb_old * (Poly.comb_old %*% b_new))

  ## update a
  xa_old <- c(x.comb_old %*% a_old)
  if(positive) xa_old <- pmax(0, xa_old)
  xa_new <- c(x.comb_old %*% a_new)
  if(positive) xa_new <- pmax(0, xa_new)
  
  pg.residual_old <- pg.residual_old + temp * xa_old * Half.comb_old
  res_new <- pg.residual_old - temp * xa_new * Half.comb_old
  log.likelihood <- sum(dnorm(res_new,0,sqrt(1/pg.draws_old),log=TRUE))
  
  return(log.likelihood)
}

## likelihood ratio

for(l in 1:(L-1)){
  config_info1 <- config_info_all[[l]]
  config_info2 <- config_info_all[[l+1]]
  
  loglike12 <- get.swap.likelihood(temp=temp_PT[l],
                                   pg.residual_old=config_info1$pg.residual,
                                   x.comb_old=config_info1$x.comb,
                                   Half.comb_old=config_info1$Half.comb,
                                   pg.draws_old=config_info1$pg.draws,
                                   Poly.comb_old=config_info1$Poly.comb,
                                   b_old=b_PT[,,l], b_new=b_PT[,,(l+1)],
                                   a_old=a_PT[,l], a_new=a_PT[,(l+1)], positive)
  loglike21 <- get.swap.likelihood(temp=temp_PT[l+1],
                                   pg.residual_old=config_info2$pg.residual,
                                   x.comb_old=config_info2$x.comb,
                                   Half.comb_old=config_info2$Half.comb,
                                   pg.draws_old=config_info2$pg.draws,
                                   Poly.comb_old=config_info2$Poly.comb,
                                   b_old=b_PT[,,(l+1)], b_new=b_PT[,,l],
                                   a_old=a_PT[,(l+1)], a_new=a_PT[,l], positive)
  
  loglikeratio <- (loglike12 + loglike21 - log.likelihood_PT[l] - log.likelihood_PT[l+1])
  swap_accept_prob[l] <- min(1, exp(loglikeratio))
}
```



# data

```{r}
vote <- read.csv("./codes/vote_data.csv")
str(vote)
vote <- subset(vote, abs(from_requirement_threshold) < 0.5)
set.seed(1234)
train.ss <- sample(nrow(vote), 15e3)

preds <- model.matrix(~ #major.requirement 
                        + ISS.against.recommendation 
                      #+ shares.oustanding.base 
                      #+ special.meeting 
                      + analyst.coverage 
                      #+ institutional.ownership 
                      + past.stock.return 
                      + q 
                      + firm.size, 
                      data = vote)[,-1]
preds.train <- scale(preds[train.ss,])
x <- cbind(1, preds.train)
x.test <- cbind(1, scale(preds[-train.ss,],
                         center=attr(preds.train,"scaled:center"),
                         scale=attr(preds.train,"scaled:scale")))
dimnames(x)[[2]][1] <- "Intercept"
dimnames(x.test)[[2]][1] <- "Intercept"
x.names <- dimnames(x)[[2]]

y <- 2*vote[train.ss,"from_requirement_threshold"]
y.test <- 2*vote[-train.ss,"from_requirement_threshold"]

bunch <- FALSE
pos.est <- TRUE
```

```{r}
set.seed(123)
N_para <- 6
p <- 6
order <- 15
L <- 5

kbw <- 2/(order-1)
knots <- seq(-1,1,kbw)
bsd.sq <- (0.67*kbw)^2
gausskern <- lapply(knots, function(mu) return(function(x) return(exp(-0.5*(x-mu)^2/bsd.sq))))
get.poly.mat <- function(y) return(sapply(gausskern, function(f) f(y)))

yFn <- get.poly.mat

b_init <- replicate(N_para, matrix(rnorm(n = order*p, sd=5), order, p))
a_init <- replicate(N_para, rnorm(n = p, sd = 1.5))
```


# bdregjump_adapt_tempered_new

```{r eval=FALSE}
adapt_tempered_new <- foreach(i = 1:N_para) %dorng%
  bdregjump_adapt_tempered(y=y, x=x, b=b_init[,,i], burn=0, nsamp=5000, thin=1, order = 15,
                     jump=list(a=a_init[,i], prec = 1, positive=pos.est,
                               persistence=0.68, update.jbw=FALSE),
                     adap=list(parallel = 5, alpha_star_PT = 0.234))
```

```{r}
load("./report/report14/adapt_tempered_new.RData")
```

```{r temperednew, cache=TRUE}
adapt_tempered_new_coda <- multi_chain_parallel_coda(adapt_tempered_new, burn=0, N_para, p, order)

plot(adapt_tempered_new_coda)
```

```{r}
sapply(1:6, function(x) adapt_tempered_new[[x]]$acpt)[c(1,3),]
```

```{r}
get.swap.ratio(adapt_tempered_new)
```

```{r}
para_temp_plot(adapt_tempered_new, 5, N_para)
```

I also have tried set inverse temp to be (1, 1, 1, 1, 1) at the beginning without
adaption of temperature. However, the swap prob is still 0.

# bdregjump_tempered_adapt

## temp = 0.1

```{r eval=FALSE}
tempered_adapt01 <- foreach(i = 1:N_para) %dorng%
  bdregjump_tempered_adapt(y=y, x=x, b=b_init[,,i], burn=0, nsamp=5000, thin=1,
                           order = 15, temp = 0.1,
                     jump=list(a=a_init[,i], prec = 1, positive=pos.est,
                               persistence=0.68, update.jbw=FALSE))
```

```{r}
load("./report/report14/tempered_adapt01.RData")
```

```{r adapt01, cache=TRUE}
tempered_adapt01_coda <- multi_chain_coda(tempered_adapt01, burn=0, N_para, p, order)

plot(tempered_adapt01_coda)
```


## temp = 0.25

```{r eval=FALSE}
tempered_adapt025 <- foreach(i = 1:N_para) %dorng%
  bdregjump_tempered_adapt(y=y, x=x, b=b_init[,,i], burn=0, nsamp=5000, thin=1,
                           order = 15, temp = 0.25,
                     jump=list(a=a_init[,i], prec = 1, positive=pos.est,
                               persistence=0.68, update.jbw=FALSE))
```

```{r}
load("./report/report14/tempered_adapt025.RData")
```

```{r adapt025, cache=TRUE}
tempered_adapt025_coda <- multi_chain_coda(tempered_adapt025, burn=0, N_para, p, order)

plot(tempered_adapt025_coda)
```

## temp = 0.35

```{r eval=FALSE}
tempered_adapt035 <- foreach(i = 1:N_para) %dorng%
  bdregjump_tempered_adapt(y=y, x=x, b=b_init[,,i], burn=0, nsamp=5000, thin=1,
                           order = 15, temp = 0.35,
                     jump=list(a=a_init[,i], prec = 1, positive=pos.est,
                               persistence=0.68, update.jbw=FALSE))
```

```{r}
load("./report/report14/tempered_adapt035.RData")
```

```{r adapt035, cache=TRUE}
tempered_adapt035_coda <- multi_chain_coda(tempered_adapt035, burn=0, N_para, p, order)

plot(tempered_adapt035_coda)
```

## temp = 0.5

```{r eval=FALSE}
tempered_adapt05 <- foreach(i = 1:N_para) %dorng%
  bdregjump_tempered_adapt(y=y, x=x, b=b_init[,,i], burn=0, nsamp=5000, thin=1,
                           order = 15, temp = 0.5,
                     jump=list(a=a_init[,i], prec = 1, positive=pos.est,
                               persistence=0.68, update.jbw=FALSE))
```

```{r}
load("./report/report14/tempered_adapt05.RData")
```

```{r adapt05, cache=TRUE}
tempered_adapt05_coda <- multi_chain_coda(tempered_adapt05, burn=0, N_para, p, order)

plot(tempered_adapt05_coda)
```


