---
title: "report15"
author: "Haoliang Zheng"
date: "2/22/2022"
output:
  pdf_document:
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "/Users/zhl/Desktop/drdisc")
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.align = 'center', 
                      fig.width=6, fig.height=7)

library(abind)
library(knitr)
library(latex2exp)
require(magrittr)
require(plyr)
library(tidyverse)
library(coda)
library(doParallel)
library(doMC) 
library(foreach)
library(doRNG)

adapt <- FALSE
bunch <- FALSE
N_para <- 6

registerDoMC(6)

par(mar = c(2,2,2,2))
```


```{r source, include=FALSE}
source("./codes/new-codes.R")
source("./codes/test-codes.R")
source("./codes/other-codes.R")

multi_chain_coda_noa <- function(result_multi, burn = 1000, N_para, p, order){
  coda_result <- list()
  matrix_result <- NULL
  for(i in 1:N_para){
    matrix_result <- NULL
    for(j in 1:order){
      for(k in 1:p){
        bijk <- matrix(result_multi[[i]]$b[j,k,], ncol = 1)
        colnames(bijk) <- paste0("b",j,k)
        matrix_result <- cbind(matrix_result, bijk)
      }
    }
    coda_result[[i]] <- mcmc(matrix_result, start = burn+1)
  }
  coda_result_list <- mcmc.list(coda_result)
  
  return(coda_result_list)
}

para_temp_plot <- function(adapt_PT_new, L, N_para = 6){
par(mar = c(2,2,2,2))
par(mfrow = c(N_para, 1))
for(i in 1:N_para){
  plot(adapt_PT_new[[i]]$temp[1,], type = "l", ylim=range(adapt_PT_new[[i]]$temp),
       main = paste0("starting point ", i))
  for(j in 2:L){
    lines(adapt_PT_new[[i]]$temp[j,], col=tcol(j,.5), lwd=2)
    }
  }
}
```

# data

```{r}
vote <- read.csv("./codes/vote_data.csv")
vote <- subset(vote, abs(from_requirement_threshold) < 0.5)
set.seed(1234)
train.ss <- sample(nrow(vote), 15e3)

preds <- model.matrix(~ #major.requirement 
                        + ISS.against.recommendation 
                      #+ shares.oustanding.base 
                      #+ special.meeting 
                      + analyst.coverage 
                      #+ institutional.ownership 
                      + past.stock.return 
                      + q 
                      + firm.size, 
                      data = vote)[,-1]
preds.train <- scale(preds[train.ss,])
x <- cbind(1, preds.train)
x.test <- cbind(1, scale(preds[-train.ss,],
                         center=attr(preds.train,"scaled:center"),
                         scale=attr(preds.train,"scaled:scale")))
dimnames(x)[[2]][1] <- "Intercept"
dimnames(x.test)[[2]][1] <- "Intercept"
x.names <- dimnames(x)[[2]]

y <- 2*vote[train.ss,"from_requirement_threshold"]
y.test <- 2*vote[-train.ss,"from_requirement_threshold"]

bunch <- FALSE
pos.est <- TRUE
```

```{r}
set.seed(123)
N_para <- 6
p <- 6
order <- 15
L <- 5

kbw <- 2/(order-1)
knots <- seq(-1,1,kbw)
bsd.sq <- (0.67*kbw)^2
gausskern <- lapply(knots, function(mu) return(function(x) return(exp(-0.5*(x-mu)^2/bsd.sq))))
get.poly.mat <- function(y) return(sapply(gausskern, function(f) f(y)))

yFn <- get.poly.mat

b_init <- replicate(N_para, matrix(rnorm(n = order*p, sd=5), order, p))
a_init <- replicate(N_para, rnorm(n = p, sd = 1.5))
```


Denote $y^*_l$ as the missing history for chain $l$, then swap probability
for chain 1 and 2 is:
$$
\alpha(1, 2) = 1 \wedge \frac{p_1(y, y^*_2 | \theta_2) * p_2(y, y^*_1 | \theta_1)}{p_1(y, y^*_1 | \theta_1) * p_2(y, y^*_2 | \theta_2)}
$$

```{r eval=FALSE}
adapt_tempered_joint_new_per <- foreach(i = 1:N_para) %dorng%
  bdregjump_adapt_tempered(y=y, x=x, b=b_init[,,i], burn=0, nsamp=10, thin=1, order = 15,
                     jump=list(a=a_init[,i], prec = 1, positive=pos.est,
                               persistence=0.91, update.jbw=FALSE),
                     adap=list(parallel = 5, alpha_star_PT = 0.234))
```

```{r}
sapply(1:6, function(x) adapt_tempered_joint_new_per[[x]]$acpt)
```


```{r temperednew, cache=TRUE}
adapt_tempered_joint_new_per_coda <- multi_chain_parallel_coda(adapt_tempered_joint_new_per, burn=0, N_para, p, order)

plot(adapt_tempered_joint_new_per_coda)
```

```{r}
para_temp_plot(adapt_tempered_joint_new_per, 5, N_para)
```

# bdregjump_adapt_tempered_new

```{r eval=FALSE}
adapt_tempered_joint_new <- foreach(i = 1:N_para) %dorng%
  bdregjump_adapt_tempered(y=y, x=x, b=b_init[,,i], burn=0, nsamp=5000, thin=1, order = 15,
                     jump=list(a=a_init[,i], prec = 1, positive=pos.est,
                               persistence=0.68, update.jbw=FALSE),
                     adap=list(parallel = 5, alpha_star_PT = 0.234))
```

```{r}
load(file = "./report/report15/adapt_tempered_joint_new.RData")
```


```{r}
par(mfrow = c(3,2))
for(k in 1:6){
  plot(adapt_tempered_joint_new[[k]]$a_PT[2,1,1:5000], type = "l")
for(i in 2:5){
  lines(adapt_tempered_joint_new[[k]]$a_PT[2,i,1:5000], col = tcol(i,.5))
}
}
```


## Trace plot

```{r temperednew, cache=TRUE}
adapt_tempered_joint_new_coda <- multi_chain_parallel_coda(adapt_tempered_joint_new,
                                                           burn=0, N_para, p, order,
                                                           moreburn = 1000)

plot(adapt_tempered_joint_new_coda)

gelman.diag(adapt_tempered_joint_new_coda)

adapt_tempered_joint_new[[1]]$a_PT[1,1,1001:5000]

a_1_all <- c()
for(k in 1:6){
  a_1_all <- c(a_1_all, adapt_tempered_joint_new[[k]]$a_PT[1,1,1001:5000])
}
HPD(a_1_all)

a_2_all <- c()
for(k in 1:6){
  a_1_all <- c(a_1_all, adapt_tempered_joint_new[[k]]$a_PT[1,1,1001:5000])
}


a_all_df <- NULL

for(j in 1:6){
  a_1_all <- c()
  for(k in 1:6){
  a_1_all <- c(a_1_all, adapt_tempered_joint_new[[k]]$a_PT[j,1,1001:5000])
  }
  a_all_df <- cbind(a_all_df, a_1_all)
}

apply(a_all_df, 2, HPD)

colnames(a_all_df) <- paste0("a",1:6)
a_all_mcmc <- mcmc(a_all_df)
HPDinterval(a_all_mcmc)
HPDinterval(a_all_mcmc, prob = 0.9)

dens1 <- density(a_all_df[,1])
plot(dens1)
dens1$x[which.max(dens1$y)]
hdi(dens1, allowSplit=TRUE, credMass = 0.9)

dens2 <- density(a_all_df[,2])
plot(dens2)
dens2$x[which.max(dens2$y)]
hdi(dens2, allowSplit=TRUE, credMass = 0.9)

dens3 <- density(a_all_df[,3])
plot(dens3)
dens3$x[which.max(dens3$y)]
hdi(dens3, allowSplit=TRUE, credMass = 0.9)

dens4 <- density(a_all_df[,4])
plot(dens4)
dens4$x[which.max(dens4$y)]
hdi(dens4, allowSplit=TRUE, credMass = 0.9)

dens5 <- density(a_all_df[,5])
plot(dens5)
dens5$x[which.max(dens5$y)]
hdi(dens5, allowSplit=TRUE, credMass = 0.9)

dens6 <- density(a_all_df[,6])
plot(dens6)
dens6$x[which.max(dens6$y)]
hdi(dens6, allowSplit=TRUE, credMass = 0.9)

hdi(a_all_mcmc[,1], allowSplit=TRUE)
hdi(a_all_mcmc[,2], allowSplit=TRUE)
hdi(a_all_mcmc[,3], allowSplit=TRUE)
hdi(a_all_mcmc[,4], allowSplit=TRUE)
hdi(a_all_mcmc[,5], allowSplit=TRUE)
hdi(a_all_mcmc[,6], allowSplit=TRUE)
```

From the trace plot, we can see clear sign of swapping.

## Acceptance rate

Overall acceptance rate of Metropolis and swap for 6 different initializing points:

```{r}
sapply(1:6, function(x) adapt_tempered_joint_new[[x]]$acpt)[c(1,3),]
```

Average acceptance rate of swap for each pair of chains from 6 different initializing points:

```{r}
get.swap.ratio(adapt_tempered_joint_new)
```

According to the acceptance rate, we can see the result seems to be good.
The acceptance rate of swap is around the setting of 0.234.

I also check the last 10 computed acceptance probabilities from 6 different initializing points
to make sure that they are not taking values only 0 or 1.


```{r}
adapt_tempered_joint_new[[1]]$swap_accept_prob[,4991:5000]
adapt_tempered_joint_new[[2]]$swap_accept_prob[,4991:5000]
adapt_tempered_joint_new[[3]]$swap_accept_prob[,4991:5000]
adapt_tempered_joint_new[[4]]$swap_accept_prob[,4991:5000]
adapt_tempered_joint_new[[5]]$swap_accept_prob[,4991:5000]
adapt_tempered_joint_new[[6]]$swap_accept_prob[,4991:5000]
```

## Inverse temperature

One thing not ideal is that other four inverse temperatures are all close to 1.
However, the good thing is that they are not extremely close to 1. It seems
that they converge to 0.98, 0.96, 0.94 and 0.92.

```{r}
para_temp_plot(adapt_tempered_joint_new, 5, N_para)
```

```{r}
sapply(1:6, function(x) adapt_tempered_joint_new[[x]]$temp[,5000])
```


# order = 5, parallel = 10

```{r}
set.seed(123)
N_para <- 6
p <- 6
order <- 5
L <- 10

kbw <- 2/(order-1)
knots <- seq(-1,1,kbw)
bsd.sq <- (0.67*kbw)^2
gausskern <- lapply(knots, function(mu) return(function(x) return(exp(-0.5*(x-mu)^2/bsd.sq))))
get.poly.mat <- function(y) return(sapply(gausskern, function(f) f(y)))

yFn <- get.poly.mat

b_init <- replicate(N_para, 
                    cbind(rnorm(n=5, sd=3), rnorm(n=5, sd=1.5), rnorm(n=5, sd=0.5),
                          rnorm(n=5, sd=0.25), rnorm(n=5, sd=0.15), rnorm(n=5, sd=0.5))
                    )
a_init <- replicate(N_para, rnorm(n = p, sd = 1.5))
```


```{r eval=FALSE}
i =1 
adapt_tempered_o5_p20 <-
  bdregjump_adapt_tempered(y=y, x=x, b=b_init[,,i], burn=0, nsamp=500, thin=1, order = 5,
                     jump=list(a=a_init[,i], prec = 1, positive=pos.est,
                               persistence=0.68, update.jbw=FALSE),
                     adap=list(parallel = 20, alpha_star_PT = 0.234),
                     print.process = TRUE, print.temp = TRUE)
```

```{r}
save(adapt_tempered_o5_p20, y, x, b_init, a_init, order,
     file = "./report/report15/adapt_tempered_o5_p20.RData")
```

```{r}
par(mfrow = c(3,2))
for(i in 1:15){
  for(j in 1:6){
    plot(adapt_tempered_o5_p10$b[i,j,1:500], type = "l")
  }
}
```

```{r}
for(i in 1:6){
  plot(adapt_tempered_o5_p10$a_PT[i,1,1:500], type = "l")
}
```


```{r eval=FALSE}
i =1 
adapt_tempered_o5_p20_alpha01 <-
  bdregjump_adapt_tempered(y=y, x=x, b=b_init[,,i], burn=0, nsamp=500, thin=1, order = 5,
                     jump=list(a=a_init[,i], prec = 1, positive=pos.est,
                               persistence=0.68, update.jbw=FALSE),
                     adap=list(parallel = 20, rho_adap = -2, alpha_star_PT = 0.1),
                     print.process = TRUE, print.temp = TRUE)
```


```{r temperednew, cache=TRUE}
adapt_tempered_joint_new_alpha01_coda <- multi_chain_parallel_coda(adapt_tempered_joint_new_alpha01, burn=0, N_para, p, order)

plot(adapt_tempered_joint_new_alpha01_coda)
```

```{r}
sapply(1:6, function(x) adapt_tempered_joint_new_alpha01[[x]]$acpt)[c(1,3),]
```

```{r}
get.swap.ratio(adapt_tempered_joint_new_alpha01)
```

```{r}
para_temp_plot(adapt_tempered_joint_new_alpha01, 5, N_para)
```

```{r}
sapply(1:6, function(x) adapt_tempered_joint_new_alpha01[[x]]$temp[,5000])
```


```{r eval=FALSE}
i =1 
adapt_tempered_fix <-
  bdregjump_adapt_tempered_fix(y=y, x=x, b=b_init[,,i], burn=0, nsamp=500, thin=1, order = 5,
                     jump=list(a=a_init[,i], prec = 1, positive=pos.est,
                               persistence=0.68, update.jbw=FALSE),
                     adap=list(parallel = 10, temp_end = 0.01),
                     print.process = TRUE, print.swap = TRUE)
```

```{r}
par(mfrow = c(3,2))
for(i in 1:5){
  for(j in 1:6){
    plot(adapt_tempered_fix$b[i,j,1:500], type = "l")
  }
}
```

```{r}
for(i in 1:6){
  plot(adapt_tempered_fix$a_PT[i,1,1:500], type = "l")
  for(l in 2:L){
    lines(adapt_tempered_fix$a_PT[i,l,1:500], col=tcol(l,.5))
  }
}
```


```{r}
par(mfrow = c(3,2))
for(i in 1:5){
  for(j in 1:6){
    plot(adapt_tempered_o5_p20_alpha01$b[i,j,1:500], type = "l")
  }
}
```

```{r}
for(i in 1:6){
  plot(adapt_tempered_o5_p20_alpha01$a_PT[i,1,1:500], type = "l")
  for(l in 2:L){
    lines(adapt_tempered_o5_p20_alpha01$a_PT[i,l,1:500], col=tcol(l,.5))
  }
}
```


## one temp = 0.5

```{r eval=FALSE}
adapt_onetemp05 <- foreach(i = 1:N_para) %dorng%
  bdregjump_adapt_onetemp(y=y, x=x, b=b_init[,,i], burn=0, nsamp=5000, thin=1,
                           order = 5, temp = 0.5,
                     jump=list(a=a_init[,i], prec = 1, positive=pos.est,
                               persistence=0.68, update.jbw=FALSE))
```


```{r}
save(adapt_onetemp05, y, x, b_init, a_init, order,
     file = "./report/report15/adapt_onetemp05.RData")
```

```{r onetemp05, cache=TRUE}
adapt_onetemp05_coda <- multi_chain_coda(adapt_onetemp05, burn=0, N_para, p, order)

plot(adapt_onetemp05_coda)
```


## one temp = 0.1

```{r eval=FALSE}
adapt_onetemp01 <- foreach(i = 1:N_para) %dorng%
  bdregjump_adapt_onetemp(y=y, x=x, b=b_init[,,i], burn=0, nsamp=5000, thin=1,
                           order = 5, temp = 0.1,
                     jump=list(a=a_init[,i], prec = 1, positive=pos.est,
                               persistence=0.68, update.jbw=FALSE))
```


```{r}
save(adapt_onetemp01, y, x, b_init, a_init, order,
     file = "./report/report15/adapt_onetemp01.RData")
```

```{r onetemp05, cache=TRUE}
adapt_onetemp01_coda <- multi_chain_coda(adapt_onetemp01, burn=0, N_para, p, order)

plot(adapt_onetemp01_coda)
```

## one temp = 0.25

```{r eval=FALSE}
adapt_onetemp025 <- foreach(i = 1:N_para) %dorng%
  bdregjump_adapt_onetemp(y=y, x=x, b=b_init[,,i], burn=0, nsamp=5000, thin=1,
                           order = 5, temp = 0.25,
                     jump=list(a=a_init[,i], prec = 1, positive=pos.est,
                               persistence=0.68, update.jbw=FALSE))
```


```{r}
save(adapt_onetemp025, y, x, b_init, a_init, order,
     file = "./report/report15/adapt_onetemp025.RData")
```

```{r onetemp05, cache=TRUE}
adapt_onetemp025_coda <- multi_chain_coda(adapt_onetemp025, burn=0, N_para, p, order)

plot(adapt_onetemp025_coda)
```

## one temp = 0.01

```{r eval=FALSE}
adapt_onetemp001 <- foreach(i = 1:N_para) %dorng%
  bdregjump_adapt_onetemp(y=y, x=x, b=b_init[,,i], burn=0, nsamp=5000, thin=1,
                           order = 5, temp = 0.01,
                     jump=list(a=a_init[,i], prec = 1, positive=pos.est,
                               persistence=0.68, update.jbw=FALSE))
```


```{r}
save(adapt_onetemp001, y, x, b_init, a_init, order,
     file = "./report/report15/adapt_onetemp001.RData")
```

```{r onetemp05, cache=TRUE}
adapt_onetemp001_coda <- multi_chain_coda(adapt_onetemp001, burn=0, N_para, p, order)

plot(adapt_onetemp001_coda)
```



# order = 15

## one temp = 0.25

```{r eval=FALSE}
adapt_onetemp025_o15 <- foreach(i = 1:N_para) %dorng%
  bdregjump_adapt_onetemp(y=y, x=x, b=b_init[,,i], burn=0, nsamp=5000, thin=1,
                           order = 15, temp = 0.25,
                     jump=list(a=a_init[,i], prec = 1, positive=pos.est,
                               persistence=0.68, update.jbw=FALSE))
```


```{r}
save(adapt_onetemp025_o15, y, x, b_init, a_init, order,
     file = "./report/report15/adapt_onetemp025_o15.RData")
```

```{r onetemp05, cache=TRUE}
adapt_onetemp025_o15_coda <- multi_chain_coda(adapt_onetemp025_o15, burn=0, N_para, p, order)

plot(adapt_onetemp025_o15_coda)
```


## one temp = 0.1

```{r eval=FALSE}
adapt_onetemp01_o15 <- foreach(i = 1:N_para) %dorng%
  bdregjump_adapt_onetemp(y=y, x=x, b=b_init[,,i], burn=0, nsamp=5000, thin=1,
                           order = 15, temp = 0.1,
                     jump=list(a=a_init[,i], prec = 1, positive=pos.est,
                               persistence=0.68, update.jbw=FALSE))
```


```{r}
save(adapt_onetemp01_o15, y, x, b_init, a_init, order,
     file = "./report/report15/adapt_onetemp01_o15.RData")
```

```{r onetemp05, cache=TRUE}
adapt_onetemp01_o15_coda <- multi_chain_coda(adapt_onetemp01_o15, burn=0, N_para, p, order)

plot(adapt_onetemp01_o15_coda)
```


```{r eval=FALSE}
adapt_tempered_fix_multic <- foreach(i = 1:N_para) %dorng%
  bdregjump_adapt_tempered_fix(y=y, x=x, b=b_init[,,i], burn=0, nsamp=4000,
                               thin=1, order = 5,
                     jump=list(a=a_init[,i], prec = 1, positive=pos.est,
                               persistence=0.68, update.jbw=FALSE),
                     adap=list(parallel = 20,
                               temp_end = 0.01),
                     print.process = TRUE, print.swap = TRUE)
```


```{r}
save(adapt_tempered_fix_multic, y, x, b_init, a_init, order,
     file = "./report/report15/adapt_tempered_fix_multic.RData")
```

```{r onetemp05, cache=TRUE}
adapt_tempered_fix_multic_coda <- multi_chain_parallel_coda(adapt_tempered_fix_multic, burn=0, N_para, p, order)

plot(adapt_tempered_fix_multic_coda)
```

```{r}
get.swap.ratio(adapt_tempered_fix_multic, L = 20)
```



```{r eval=FALSE}
adapt_tempered_fix_multic2 <- foreach(i = 1:N_para) %dorng%
  bdregjump_adapt_tempered_fix(y=y, x=x, b=b_init[,,i], burn=0, nsamp=3500,
                               thin=1, order = 15,
                     jump=list(a=a_init[,i], prec = 1, positive=pos.est,
                               persistence=0.68, update.jbw=FALSE),
                     adap=list(parallel = 20,
                               temp_end = 0.01),
                     print.process = TRUE, print.swap = TRUE)
```

```{r}
save(adapt_tempered_fix_multic2, y, x, b_init, a_init, order,
     file = "./report/report15/adapt_tempered_fix_multic2.RData")
```

```{r onetemp05, cache=TRUE}
adapt_tempered_fix_multic2_coda <- multi_chain_parallel_coda(adapt_tempered_fix_multic2, burn=0, N_para, p, order)

plot(adapt_tempered_fix_multic2_coda)
```

```{r}
get.swap.ratio(adapt_tempered_fix_multic2, L = 20)
```


```{r}
par(mfrow = c(3,2))
for(i in 1:6){
  for(k in 1:6){
  plot(adapt_tempered_fix_multic2[[k]]$a_PT[i,20,1:3500], type = "l")
  lines(adapt_tempered_fix_multic2[[k]]$a_PT[i,19,1:3500], col = tcol(i,.5))
}
}

```