---
title: "report13"
author: "Haoliang Zheng"
date: "2/15/2022"
output:
  pdf_document:
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "/Users/zhl/Desktop/drdisc")
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.align = 'center', 
                      fig.width=6, fig.height=7)

library(abind)
library(knitr)
library(latex2exp)
require(magrittr)
require(plyr)
library(tidyverse)
library(coda)
library(doParallel)
library(doMC) 
library(foreach)
library(doRNG)

adapt <- FALSE
bunch <- FALSE
N_para <- 6

registerDoMC(6)

par(mar = c(2,2,2,2))
```

The posterior distribution:
$$
\pi(a, \beta) = k^{-1} p(x | a)^\beta p(a)
$$

For Simulated Tempering, we are considering $(a, \beta_1) \rightarrow (a, \beta_2)$, then
acceptance probability is:
$$
\alpha(1, 2) = 1 \wedge \frac{p(x | a)^{\beta_2} / c(\beta_2)}{p(x | a)^{\beta_1} / c(\beta_1)}
$$

For Parallel Tempering, we are considering $(a_1, \beta_1) (a_2, \beta_2) \rightarrow (a_2, \beta_1) (a_1, \beta_2)$, then swap probability is:
$$
\begin{aligned}
\alpha(1, 2) &= 1 \wedge \frac{p(x | a_1)^{\beta_2} / c(\beta_2)}{p(x | a_1)^{\beta_1} / c(\beta_1)}
\times \frac{p(x | a_2)^{\beta_1} / c(\beta_1)}{p(x | a_2)^{\beta_2} / c(\beta_2)} \\
&= 1 \wedge \left(\frac{p(x | a_2)}{p(x | a_1)}\right)^{\beta_1 - \beta_2}
\end{aligned}
$$

In our case, we are considering $((a_1,b_1), \beta_1) ((a_2,b_2), \beta_2) \rightarrow ((a_2,b_2), \beta_1) ((a_1,b_1), \beta_2)$, then swap probability should still be:
$$
\alpha(1, 2) = 1 \wedge \left(\frac{p(x | a_2, b_2)}{p(x | a_1, b_1)}\right)^{\beta_1 - \beta_2}
$$

Because we are computing likelihood with the following codes, this should be fine.
```{r eval=FALSE}
res <- pg.residual - xa * Half.comb
log.likelihood <- sum(dnorm(res,0,sqrt(1/pg.draws),log=TRUE))
```

However, another question is: our model has the following form:
$$
f(y|x) = \frac{\Phi \big(\sum_{k=1}^K (x^T\beta_k) P_k(y) - (x^T \alpha)_{+} Q(y; \lambda) I(y > 0)\big)}{\int_{-1}^{1} \Phi \big(\sum_{k=1}^K (x^T\beta_k) P_k(y) - (x^T \alpha)_{+} Q(y; \lambda) I(y > 0)\big) \mathrm{d}t}, 
\quad x \in \mathbb{R}^p, y \in (-1, 1),
$$

In `get.miss.reject.x()`, we have
```{r eval=FALSE}
w <- rowSums(x.remain * (pm %*% b)) + xa.remain*hk
u <- (runif(n.remain) < Phi(w))
w.miss <- c(w.miss, w[!u])
```
and
```{r eval=FALSE}
w.obs <- rowSums(x * (Poly.obs %*% b)) + xa.obs * Half.obs
w.miss <- missing.stuff$w.miss
w.comb <- c(w.obs, w.miss)
pg.draws <- pmax(1e-12, rpg.devroye(Ntot, h=1, z=w.comb))
pg.residual <- pg.resp - w.comb
```

Before, I only used `temp` when computing log.hastings.ratio for $a$ parameter,
```{r eval=FALSE}
log.hastings.ratio <- (temp * ( sum(dnorm(res.new,0,sqrt(1/pg.draws),log=TRUE)
                                    - dnorm(res,0,sqrt(1/pg.draws),log=TRUE)) )
                       + sum(dnorm(a.new,0,sqrt(1/hprec),log=TRUE)
                             - dnorm(a,0,sqrt(1/hprec),log=TRUE)))
```

Should we add `temp` in the `get.miss.reject.x()` and `w.obs` to reflect
different temperature configurations?




$$
\alpha(1, 2) = 1 \wedge \frac{p_1(y, y^*_2 | \theta_2) * p_2(y, y^*_1 | \theta_1)}{p_1(y, y^*_1 | \theta_1) * p_2(y, y^*_2 | \theta_2)}
$$